Characterizing the Human Virome and Colorectal Cancer Link Using Machine Learning and Network Theory

### 2017-02-24 doc/manuscript.Rmd
General document push. In this case, more data figures.

@@

>>>>

## Phage Replication Style Identification
Phage OGU replication sytle was identified using methods described previously [@Minot:2011ez; @Hannigan:2015fz; @Hannigan:2017ky]. Briefly, we identified lysogenic phage OGUs as representative contigs containing at least one of three genomic markers: 1) phage integrase genes, 2) prophage genes from the ACLAME database, 3) genomic similarity to bacterial reference genomes. Integrase genes were identified in phage OGU representative contigs by aligning the contigs to a reference database of all known phage integrase genes from the Uniprot database (Uniprot search term: "organism:phage gene:int NOT putative"). Prophage genes were identified in the same way, using the ACLAME set of reference prophage genes. In both cases, the blastx algorithm was used with an e-value of 10e-5. Representative contigs were also identified as potential lysogenic phages by having a high genomic similarity to bacterial genomes. To accomplish this, representative phage contigs were aligned to the European Nucleotide Archive bacterial genome reference set using the blastn algorithm (e-value < 10e-25).

@@

![*Lysogenic phage relative abundance in disease states. Phage OGUs were predicted to be either lytic or lysogenic, and the relative abundance of lysogenic phages was quantified and represented as a boxplot. No disease groups were statistically significant.*\label{betaogunegative}](../figures/lysogenic-relabund.pdf){ width=50% }

![*Relative abundance correlations between bacterial OTUs and virome OGUs. A) Pearson correlation R values between all bacterial OTUs (x-axis) and viral OGUs (y-axis) with blue being positively correlated and red being negatively correlated. Operational units are organized by importance in their colorectal cancer classification models, such that the most important units are in the top left corner. B) Magnification of the boxed region in pannel (A), highlighting the correlation between the most important bacterial OTUs and virome OGUs. The most important operational units are in the top left corner of the heatmap, and the correlation scale is the same as pannel (A). C) Histogram quantifying the frequencies of Pearson correlation coefficients between all bacterial OTUs and virome OGUs.*\label{betaogunegative}](../figures/phage-bacteria-cor.pdf){ width=100% }
